opponent modelling

### 2020-01-22 comp_model_1019.Rmd
report changes to comp modelling section

## Modelling fixed level-k type participant
```{r}
# FIXED STRATEGIES (level-k) 
# create a function which computes the -2 log Likelihood over the three games for a particular strategy
# modelLL <- function(par,data,strategy, return_value=c("-2logLik","likelihood_by_trial")) {
#   # par = parameter vector (just has one element here) 
#   # dat = data subset for one participant
#   # strategy = c("level1","level2","level3","WSLS")
#   # return = -2logLik (default, for use in optimization) or "likelihood_by_trial" (for later plotting, etc)
#   return_value <- match.arg(return_value)
#   
#   # count the number of supplied parameters. If 1, it is a _1 model, 
#   # if 3, it is a _3 model. For the _3 model we will use unbounded
#   # optimization (transforming parameters from a logit scale back to 0 to 1)
#   n_theta <- length(par)
#   if(n_theta == 3) par <-my_logistic(par)
#   theta <- par[1]
#   # set up a vector to store the likelihood
#   lik <- rep(0.0,nrow(data))
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     pred_file <- switch(as.character(t_game),rps=rps_predict,fwg = fwg_predict, numbers = numbers_predict)
#     if(n_theta == 3) {
#       theta <- switch(as.character(t_game),rps=par[1],fwg = par[2], numbers = par[3])
#     }
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       pred <- NA
#       lik[t] <- 1/nopts
#     } else {
#       # get prediction from file
#       pred <- as.character(filter(pred_file,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
#       if(as.character(data[t,"h_action"]) == pred) {
#         lik[t] <- theta + (1-theta)/nopts
#       } else {
#         lik[t] <- (1-theta)/nopts
#       }
#     }
#     # for debugging uncomment the line below
#     # cat(as.character(pred),as.character(data[t,"h_action"]),lik[t],"\n")
#   }
#   if(return_value == "-2logLik") {
#     ret <- -2*sum(log(lik))
#     if(ret == Inf) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik)
# }
# 
# all_modelling <- list()
# for(id in unique(dat$human_id)) {
#   all_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   # level 1
#   all_modelling[[id]]$level1_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level1")
#   all_modelling[[id]]$level1_3 <- optim(my_logit(rep(all_modelling[[id]]$level1_1$minimum,3)),fn=modelLL,data=tdat,strategy="level1")
#   # level 2
#   all_modelling[[id]]$level2_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level2")
#   all_modelling[[id]]$level2_3 <- optim(my_logit(rep(all_modelling[[id]]$level2_1$minimum,3)),fn=modelLL,data=tdat,strategy="level2")
#   # level 3
#   all_modelling[[id]]$level3_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level3")
#   all_modelling[[id]]$level3_3 <- optim(my_logit(rep(all_modelling[[id]]$level3_1$minimum,3)),fn=modelLL,data=tdat,strategy="level3")
#   # WSLS
#   all_modelling[[id]]$WSLS_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="WSLS")
#   all_modelling[[id]]$WSLS_3 <- optim(my_logit(rep(all_modelling[[id]]$WSLS_1$minimum,3)),fn=modelLL,data=tdat,strategy="WSLS")
# }
# 
# save(all_modelling,file="all_modelling.RData")

```{r}
# create a table with BIC
# BIC_results <- data.frame()
# for(id in unique(dat$human_id)) {
#     BIC_results <- rbind(BIC_results,
#                          data.frame(
#                            "id" = id,
#                            "condition" = dat[dat$human_id==id,"condition"][1],
#                            "random_BIC" = -2*(100*log(1/3) + 50*log(1/5)), 
#                            "level1_1_theta" = all_modelling[[id]]$level1_1$minimum,
#                            "level1_1_BIC" = all_modelling[[id]]$level1_1$objective + 1*log(150),
#                            "level1_3_theta1" = my_logistic(all_modelling[[id]]$level1_3$par[1]),
#                            "level1_3_theta2" = my_logistic(all_modelling[[id]]$level1_3$par[2]),
#                            "level1_3_theta3" = my_logistic(all_modelling[[id]]$level1_3$par[3]),
#                            "level1_3_BIC" = all_modelling[[id]]$level1_3$value + 3*log(150),
#                            "level2_1_theta" = all_modelling[[id]]$level2_1$minimum,
#                            "level2_1_BIC" = all_modelling[[id]]$level2_1$objective + 1*log(150),
#                            "level2_3_theta1" = my_logistic(all_modelling[[id]]$level2_3$par[1]),
#                            "level2_3_theta2" = my_logistic(all_modelling[[id]]$level2_3$par[2]),
#                            "level2_3_theta3" = my_logistic(all_modelling[[id]]$level2_3$par[3]),
#                            "level2_3_BIC" = all_modelling[[id]]$level2_3$value + 3*log(150),
#                            "level3_1_theta" = all_modelling[[id]]$level3_1$minimum,
#                            "level3_1_BIC" = all_modelling[[id]]$level3_1$objective + 1*log(150),
#                            "level3_3_theta1" = my_logistic(all_modelling[[id]]$level3_3$par[1]),
#                            "level3_3_theta2" = my_logistic(all_modelling[[id]]$level3_3$par[2]),
#                            "level3_3_theta3" = my_logistic(all_modelling[[id]]$level3_3$par[3]),
#                            "level3_3_BIC" = all_modelling[[id]]$level3_3$value + 3*log(150),
#                            "WSLS_1_theta" = all_modelling[[id]]$WSLS_1$minimum,
#                            "WSLS_1_BIC" = all_modelling[[id]]$WSLS_1$objective + 1*log(150),
#                            "WSLS_3_theta1" = my_logistic(all_modelling[[id]]$WSLS_3$par[1]),
#                            "WSLS_3_theta2" = my_logistic(all_modelling[[id]]$WSLS_3$par[2]),
#                            "WSLS_3_theta3" = my_logistic(all_modelling[[id]]$WSLS_3$par[3]),
#                            "WSLS_3_BIC" = all_modelling[[id]]$WSLS_3$value + 3*log(150)
#                          ))
# }
# write.csv(BIC_results,file="BIC_result.csv",row.names = FALSE)
# kable(BIC_results)
```

# create a table with best fitting models by condition
# table(dat$condition[dat$round==1 & dat$game == "rps"],c("random","level1_1","level2_1","level3_1","WSLS_1","level1_3","level2_3","level3_3","WSLS_3")[apply(BIC_results[,c("random_BIC","level1_1_BIC","level2_1_BIC","level3_1_BIC","WSLS_1_BIC","level1_3_BIC","level2_3_BIC","level3_3_BIC","WSLS_3_BIC")],1,which.min)])
# 
# ## treat 1 theta and 3 theta versions the same
# # create a table with best fitting models by condition
# table(dat$condition[dat$round==1 & dat$game == "rps"],c("random","level1","level2","level3","WSLS","level1","level2","level3","WSLS")[apply(BIC_results[,c("random_BIC","level1_1_BIC","level2_1_BIC","level3_1_BIC","WSLS_1_BIC","level1_3_BIC","level2_3_BIC","level3_3_BIC","WSLS_3_BIC")],1,which.min)])
#

## Fixed level-k type: Likelihood of the different models by trial
This plot ks based on the three parameter versions of each strategy, and shows the average probability according to each model of participants' choices. It can be seen that, compared to the RPS game, where the fit grows slowly over trials, the fit in the later games increases much more rapidly.

```{r}
# long_likelihood_by_trial <- data.frame()
# for(id in unique(dat$human_id)) {
#   tdat <- subset(dat,human_id == id)
#   for(strategy in c("level1","level2","level3","WSLS")) {
#     long_likelihood_by_trial <- rbind(long_likelihood_by_trial,
#                                       data.frame(id = id,
#                                                  condition = dat[dat$human_id==id,"condition"][1],
#                                                  strategy = strategy,
#                                                  round = 1:50,
#                                                  game = rep(c("rps","fwg","numbers"),each=50),
#                                                  likelihood = modelLL(all_modelling[[id]][[paste(strategy,"_3",sep="")]]$par,
#                                                                       data=tdat,
#                                                                       strategy = strategy,
#                                                                       return_value = "likelihood_by_trial")
#                                       )
#                                 )
#   }
# }
# library(ggplot2)
# plot_dat <- long_likelihood_by_trial %>% group_by(condition,game,round,strategy) %>%
#       summarize(probability = mean(likelihood))
# # make sure the different games are ordered in the way they were played
# plot_dat$game <- factor(plot_dat$game,levels=c("rps","fwg","numbers"))
# ggplot(plot_dat,aes(x=round,y=probability,colour=strategy)) + geom_line() + facet_grid(game~condition)
```

@@

naive_bayes <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {

@@

    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred_opp <- NA
      lik_hum[t] <- 1/nopts
      ## KEY CODE FOR LEARNING TRANSFER !! (if no OM transferreset prior to uniform)
      if (!(opp_mod_transfer)) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

@@

# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# naive_bayes(c(0.9,0.99),data,c("level0","level1","level2"),"-2loglik",FALSE)
```

```{r}
rps_predict_opp <- read.csv("rps_predict_opp.csv")
red_opp <- filter(rps_predict_opp, rps_predict_opp$human_previous == "R" & rps_predict_opp$computer_previous == "S")

as.character(red_opp[["level0"]])
```

# Fitting a bayesian model WITH TRANSFER of opp strat distribution between games 
```{r}

TR_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  TR_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # TR_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = c(0,0), upper = c(1,1), data=tdat, opp_strategy_vec = c("level0","level1","level2") ,return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE, parallelType=1,parVar = c("best_response")))

  TR_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}
save(TR_bayes_modelling,file="TR_bayes_modelling.RData")

@@

# Fitting a bayesian model with NO TRANSFER of opp strat distribution between games 
```{r}
NT_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  NT_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # NT_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))

  NT_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(NT_bayes_modelling,file="NT_bayes_modelling.RData")

@@

# Summarise results in dataframe for both transfer and no transfer models 
```{r}
Nbayes_par_results <- data.frame()
for(id in unique(dat$human_id)) {
Nbayes_par_results <- rbind(Nbayes_par_results,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         # BICs 
                         "BIC_transfer" = TR_bayes_modelling[[id]]$value + 2*log(150),
                         "BIC_no_transfer" = NT_bayes_modelling[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = TR_bayes_modelling[[id]]$par[1],
                         "theta_no_transfer" = NT_bayes_modelling[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = TR_bayes_modelling[[id]]$par[2],
                         "eps_no_transfer" = NT_bayes_modelling[[id]]$par[2]))
}

@@

## Q-learning fit
```{r}
QL_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}
save(QL_modelling,file="QL_modelling.RData")

@@

## Fitting Q_learning with states
```{r}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# Q_learn_states(c(0.1,0.1), data, "-2loglik")

@@

## Fitting Self-Tuning EWA to data 
```{r}

@@

```
```{r}
# cat(EWA_self_modelling[["38VxtUSv_h6RR5-tAAA2"]]$optim$bestmem)
```

@@

```{r}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# EWA_states(0.1, data, "-2loglik")
# optim(10.0,fn=EWA_states,gr = NULL, data=data,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")

@@

```{r}
MBM_modelling <- list()
for(id in unique(dat$human_id)) {
  MBM_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  # MBM_modelling[[id]] <- optim(c(1,0.1),fn=QMBM,gr = NULL, data=tdat,"-2loglik", gamma =0 , lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")

@@

## Putting results together 
```{r}
load("TR_bayes_modelling.RData")
load("NT_bayes_modelling.RData")
load("QL_modelling.RData")
load("QL_states_modelling.RData")
load("EWA_modelling.RData")
load("EWA_self_modelling.RData")
load("MBM_modelling.RData")
load("EWA_states_modelling.RData")

@@

                         # Bayesian updating with/without transfer
                         "Bayes_Tr_BIC" = TR_bayes_modelling[[id]]$value + 2*log(150),
                         "Bayes_No_Tr_BIC" = NT_bayes_modelling[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = TR_bayes_modelling[[id]]$par[1],
                         "theta_no_transfer" = NT_bayes_modelling[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = TR_bayes_modelling[[id]]$par[2],
                         "eps_no_transfer" = NT_bayes_modelling[[id]]$par[2],

@@

 write.csv(Table_results,file="Table_results.csv",row.names = TRUE)
 kable(Table_results)
 barplot(Table_results)

>>>>

@@

exp1_ToM <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {

@@

    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred_opp <- NA
      lik_hum[t] <- 1/nopts
      ## KEY CODE FOR LEARNING TRANSFER !! (if no OM transfer, reset prior to uniform)
      if (!(opp_mod_transfer)) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

@@

# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# exp1_ToM(c(0.9,0.99),data,c("level0","level1","level2"),"-2loglik",FALSE)
```

```{r}
# rps_predict_opp <- read.csv("rps_predict_opp.csv")
# red_opp <- filter(rps_predict_opp, rps_predict_opp$human_previous == "R" & rps_predict_opp$computer_previous == "S")
# 
# as.character(red_opp[["level0"]])
```

# Fitting a bayesian model WITH BETWEEN GAME TRANSFER of opp strat distribution between games 
```{r, cache=TRUE}
exp1_ToM_TR <- list()
for(id in unique(dat$human_id)) {
  exp1_ToM_TR[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # exp1_ToM_TR[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0,0), upper = c(1,1), data=tdat, opp_strategy_vec = c("level0","level1","level2") ,return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE, parallelType=1,parVar = c("best_response")))

  exp1_ToM_TR[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}
save(exp1_ToM_TR,file="exp1_ToM_TR.RData")

@@

# Fitting a bayesian model with NO BETWEEN GAME TRANSFER of opp strat distribution between games 
```{r, cache =TRUE}
exp1_ToM_NT <- list()
for(id in unique(dat$human_id)) {
  exp1_ToM_NT[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # exp1_ToM_NT[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))

  exp1_ToM_NT[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(exp1_ToM_NT,file="exp1_ToM_NT.RData")

@@

# Summarise results in dataframe for both transfer and no transfer models 
```{r}
Nbayes_par_results <- data.frame()
for(id in unique(dat$human_id)) {
Nbayes_par_results <- rbind(Nbayes_par_results,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         # BICs 
                         "BIC_transfer" = exp1_ToM_TR[[id]]$value + 2*log(150),
                         "BIC_no_transfer" = exp1_ToM_NT[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = exp1_ToM_TR[[id]]$par[1],
                         "theta_no_transfer" = exp1_ToM_NT[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = exp1_ToM_TR[[id]]$par[2],
                         "eps_no_transfer" = exp1_ToM_NT[[id]]$par[2]))
}

@@

## Q-learning fit
```{r, cache=TRUE}
QL_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}
save(QL_modelling,file="QL_modelling.RData")

@@

## Fitting Q_learning with states
```{r , cache =TRUE}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# Q_learn_states(c(0.1,0.1), data, "-2loglik")

@@

## Fitting Self-Tuning EWA to data 
```{r, cache=TRUE}

@@

```{r}
# cat(EWA_self_modelling[["38VxtUSv_h6RR5-tAAA2"]]$optim$bestmem)
```

@@

```{r, cache=TRUE}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# EWA_states(0.1, data, "-2loglik")
# optim(10.0,fn=EWA_states,gr = NULL, data=data,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")

@@

```{r, cache=TRUE}
MBM_modelling <- list()
for(id in unique(dat$human_id)) {
  MBM_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  # MBM_modelling[[id]] <- optim(c(1,0.1),fn=QMBM,gr = NULL, data=tdat,"-2loglik", gamma =0 , lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")

@@

## Putting results together 
```{r}
# load("exp1_ToM_TR.RData")
# load("exp1_ToM_NT.RData")
# load("QL_modelling.RData")
# load("QL_states_modelling.RData")
# load("EWA_modelling.RData")
# load("EWA_self_modelling.RData")
# load("MBM_modelling.RData")
# load("EWA_states_modelling.RData")

@@

                         # Bayesian updating with/without transfer
                         "Bayes_Tr_BIC" = exp1_ToM_TR[[id]]$value + 2*log(150),
                         "Bayes_No_Tr_BIC" = exp1_ToM_NT[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = exp1_ToM_TR[[id]]$par[1],
                         "theta_no_transfer" = exp1_ToM_NT[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = exp1_ToM_TR[[id]]$par[2],
                         "eps_no_transfer" = exp1_ToM_NT[[id]]$par[2],

@@

 write.csv(Table_results,file="Table_results.csv",row.names = TRUE)
 #kable(Table_results)
 barplot(Table_results)
